---
title: "Data 607 Assignment 3"
author: "Jayden Jiang"
date: "2024-10-09"
output: html_document
---

## Library 
```{r}
library(tidyverse)
library(dplyr)
```

## Load Data

```{r}
children_supply <- read.csv ("https://raw.githubusercontent.com/JaydeeJan/Children-Supply/refs/heads/main/Childcare_Need___Supply__All_.csv")

head(children_supply)
tail(children_supply)
```

```{r}
glimpse(children_supply)
```
## Replacing N/A to empty values 
```{r}
children_supply[children_supply == ""] <- NA
head(children_supply,10)
tail(children_supply,10)
```
## First Normal Form (1NF) sort the data by Percent Need Met from min. to max.
#### 1. Using row order to convey information is not permitted.
#### 2. Mixing data types within the same column is not permitted.
#### 3. Having a table without a primary key is not permitted.
#### 4. Repeating groups are not permitted.
```{r}
percent_met <- children_supply %>% arrange(`Percent.Need.Met`)
filtered_percentmet <- percent_met %>% filter(`Percent.Need.Met` !=0)
head(filtered_percentmet,10)
tail(filtered_percentmet,10)
```
## Second Normal Form (2NF)
#### Each non-key attribute in the table must be dependent on the entire primary key. 
```{r}
# Geographic Table
geographic_table <- filtered_percentmet %>%
  select(`Geographic.Unit`, `Geographic.ID`, `Geographic.Name`)
head(geographic_table,10)
```
```{r}
# Childcare-related info. table
childcare_table <- filtered_percentmet %>%
  select(`Geographic.ID`, `State.Median.Income.Bracket`, `Age.Group`, `Estimated.Children.Receiving.Childcare`, `Estimate.of.Unserved`, `Percent.Need.Met`)
head(childcare_table, 10)
tail(childcare_table, 10)
```

## Fourth Normal Form (4NF)
#### The only kinds of multivalued dependency allowed in a table are multivalued dependencies on the key. 
```{r}
median_income <- unique(childcare_table[, c("Geographic.ID", "State.Median.Income.Bracket")])
income_sorted <- median_income[order(median_income$Geographic.ID), ]
head(income_sorted, 10)
tail(income_sorted, 10)
```
```{r}
age_group <- unique(childcare_table[, c("Geographic.ID", "Age.Group")])
age_sorted <- age_group[order(age_group$Geographic.ID), ]
head(age_sorted, 10)
tail(age_sorted, 10)
```
## Character Manipulation

#### 2. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}
listed_majors <- read.csv ("https://raw.githubusercontent.com/fivethirtyeight/data/refs/heads/master/college-majors/majors-list.csv")

head(listed_majors)
tail(listed_majors)

# use grepl to filter rows where the Major column contains "DATA" or "STATISTICS"
filtered_majors <- listed_majors[grepl("DATA|STATISTICS", listed_majors$Major, ignore.case = TRUE), ]

print(filtered_majors)

```

## Describe, in words, what these expressions will match:

### (.)\1\1
#### 1. (.) - will match any single character in first group
#### 2. \1 - refers to the match contained in the first parenthesis
#### 3. \1\1 - character repeated two or more times. 



### "(.)(.)\\2\\1"
#### 1. (.)(.) - will match any single character in first group and second group
#### 2. \\2 refers to the match contained in the second parenthesis
#### 3. \\1 refers to the match contained in the first parenthesis

### (..)\1
#### 1. (..) - match with two characters
#### 2. \1 - refers to the match contained in the first parenthesis

### "(.).\\1.\\1"
#### 1. (.) -  will match any single character in first group
#### 2. . - match any character
#### 3. \\1 - refers to the match contained in the first parenthesis
#### 4. . - match any character
#### 5. \\1 - refers to the match contained in the first parenthesis

### "(.)(.)(.).*\\3\\2\\1"
#### 1. (.) - will match any single character in first group
#### 2. (.) - will match any single character in second group
#### 3. (.) - will match any single character in third group
#### 4. .* - will match any sequence of characters
#### 5. \\3 - refers to the match contained in the third parenthesis
#### 6. \\2 - refers to the match contained in the second parenthesis
#### 7. \\1 - refers to the match contained in the first parenthesis



## Construct regular expressions to match words that:

#### Start and end with the same character.
#### (.)\1

#### Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
#### "(..).*\\1"

#### Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
#### "(.)\\1.*\\1"


